---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


Algorithm of Anomaly Detection with Kalman Filter is:

- Set a window size. (integer)
- Iterate fixed sized window through whole data set (94836 records)
- For each window train windows records with Kalman filter to protect next record after window.
- With predicted value x, standard deviation of window records st, mean of kalman series mn and expand factor im; calculate confidence interval. Expand factor is a hyper parameter that is used to adjust confidence interval (CI).

  CI = mn +(-) (im * sd) / sqrt(windowSize)


- If next record after window is not between confidence interval, then it is an anomaly,.
    
- Strong features of this method: It is very good for point anomalies. It is executed fast. It is accurate and efficient. Needs no extra calculation like we did in Arima and statistical calulation method. We needed to find a replacement for amomaly records in learning process and find a way to get learning process out of monotonous learning. 

- Weak features of this method: It can not find frequency based anomalies. Because method always detects anomaly with confidence interval. Broken frequencies those are in confidence interval cannot be detected.  

***********************************

Set hyperparameters and initial assignments here. Hyper parameters are window size of model, kalman filter variance, expand factor 

```{r}
library("ggplot2")
# windows size parameters, start and end size can be different if one wants to repeat experiment with different sizes 
windowSizeStart = 200
windowSizeEnd = 200

measureVariance = 0.001
variance = 1e-7
intervalMultiplier = 80 # it is used to adjust model success, expand factor

minConf = 5
maxConf = 15

standardDeviation = 0
mn = 0
diff = 0
predictedValue = 0

```


Results will be stored on this data frame, it is for multiple executions (loop).

```{r}
resultSet = data.frame(matrix(ncol = 8, nrow = 0))
colnames(resultSet) <- c("windowSize", "intervalMultiplier", "variance", "measureVariance", "tp", "fp", "tn", "fn")


```


Standardized data is loaded here. This is source data to be used in this experiment

```{r}

setwd("C:/Users/sametyazak/Desktop/ynwa/bau/2017 - Thesis/code/r/")
completeData = read.csv(file="..\\..\\data\\workspace\\real_full_standard.csv", header=TRUE, sep=",")
completeData = cbind(completeData, predicted_value = 0, conf_low = 0, conf_high = 0, predicted_anomaly = 0)
completeData$calculation_value = completeData$value


```


Function that calculates confidence interval using kalman filter with current window records.


```{r}
getConfidenceInterval <- function(windowSize, windowRecords) {
  predictionSize = windowSize
  posterior = matrix(0, nrow = predictionSize, ncol = 1)
  errorPost = matrix(0, nrow = predictionSize, ncol = 1)
  priori = matrix(0, nrow = predictionSize, ncol = 1)
  Pminus = matrix(0, nrow = predictionSize, ncol = 1)
  gain = matrix(0, nrow = predictionSize, ncol = 1)
  
  errorPost[1] = 1.0
  
  for (i in 2:predictionSize) {
    priori[i] = posterior[i-1]
    Pminus[i] = errorPost[i-1] + variance
    
    gain[i] = Pminus[i] / ( Pminus[i] + measureVariance )
    posterior[i] = priori[i] + gain[i] * (windowRecords[i, "value"] - priori[i])
    errorPost[i] = ( 1 - gain[i] ) * Pminus[i]
  }
  
  
  standardDeviation = sd(windowRecords$calculation_value)
  mn = mean(posterior)
  diff = (intervalMultiplier*standardDeviation)/sqrt(windowSize)
  predictedValue = posterior[predictionSize]
  
  if (diff < minConf) {
    diff = minConf
  } else if (diff > maxConf) {
    diff = maxConf
  }
  
  return (data.frame(mn-diff, mn+diff, predictedValue))
}


```

Prediction process:


```{r}
print("Prediction Start:")
print(date())

for (windowSize in seq(windowSizeStart, windowSizeEnd, 50)) {
  print(paste0("windowSize:", windowSize))
  tp = 0; tn = 0; fp = 0; fn = 0
  
  startIndex = (windowSize+1)
  endIndex = dim(completeData)[1]
  
  windowEndIndex = windowSize
  windowStartIndex = 1
  
  windowRecords = completeData[completeData$timestamp >= windowStartIndex & completeData$timestamp <= windowEndIndex,];
  
  for (i in startIndex:endIndex) {
    
    completeData[i,"predicted_value"] = 0
    completeData[i,"conf_low"] = 0
    completeData[i,"conf_high"] = 0
    completeData[i,"predicted_anomaly"] = 0
    
    if (i != startIndex) {
      windowRecords = rbind(windowRecords[windowRecords$timestamp > (i-windowSize-1),], completeData[i-1,])
    }
    
    windowForecast = getConfidenceInterval(windowSize, windowRecords);
    
    completeData[i,"predicted_value"] = windowForecast[1,3]
    completeData[i,"conf_low"] = windowForecast[1,1]
    completeData[i,"conf_high"] = windowForecast[1,2]
    
    predictedAnomaly = if (completeData[i,"value"] <= completeData[i,"conf_high"] & completeData[i,"value"] >= completeData[i,"conf_low"]) 0 else 1
    
    completeData[i,"predicted_anomaly"] = predictedAnomaly
    
    if (predictedAnomaly == 1) {
      completeData[i,"calculation_value"] = (completeData[i,"conf_high"] + completeData[i,"conf_low"])/2
    }
    
    if (completeData[i,"is_anomaly"] == 1 & predictedAnomaly == 1) {
      tp = tp + 1;
    } else if (completeData[i,"is_anomaly"] == 1 & predictedAnomaly == 0) {
      fn = fn + 1
    } else if (completeData[i,"is_anomaly"] == 0 & predictedAnomaly == 1) {
      fp = fp + 1
    } else if (completeData[i,"is_anomaly"] == 0 & predictedAnomaly == 0) {
      tn = tn + 1
    }
  }
  
  resultSet = rbind(resultSet, data.frame(windowSize, intervalMultiplier, variance, measureVariance, tp, fp, tn, fn))
}

print("Prediction Start:")
print(date())

```

Results with different hyper parameters.

```{r}

setwd("C:/Users/sametyazak/Desktop/ynwa/bau/2017 - Thesis/code/r/")
fullResultSet = read.csv(file="kalman_result_full.csv", header=TRUE, sep=",")

fullResultSet = cbind(fullResultSet, 
      recall = (fullResultSet$tp / (fullResultSet$tp+fullResultSet$fn)),
      precision = fullResultSet$tp / (fullResultSet$tp+fullResultSet$fp)
)

fullResultSet = cbind(fullResultSet, f1_score = 2 * (fullResultSet$precision*fullResultSet$recall) / (fullResultSet$precision+fullResultSet$recall))


fullResultSet

```

Some plots of weak part of experiment:


```{r}
blockData = completeData[completeData$timestamp <= 99000,]
plotData = blockData[blockData$timestamp <= 83000 & blockData$timestamp >= 82500,]

fpData = plotData[plotData$is_anomaly == 0 & plotData$predicted_anomaly == 1,]
fnData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 0,]
tpData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 1,]

if (dim(fnData)[1] == 0) {
  fnData = rbind(fnData, plotData[1,])
}

if (dim(fpData)[1] == 0) {
  fpData = rbind(fpData, plotData[1,])
}

if (dim(tpData)[1] == 0) {
  tpData = rbind(tpData, plotData[1,])
}

ggplot() +
  geom_line(data = plotData, aes(x = timestamp, y = value, colour = "Original")) +
  geom_line(data = plotData, aes(x = timestamp, y = conf_low, colour = "conf_low_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = conf_high, colour = "conf_high_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = predicted_value, colour = "predicted_value"))  +
  #geom_line(data = plotData, aes(x = timestamp, y = calculation_value, colour = "calculation_value"))  +
  geom_point(data = fnData, aes(x = timestamp, y = value, colour = "fn")) + 
  geom_point(data = fpData, aes(x = timestamp, y = value, colour = "fp")) + 
  geom_point(data = tpData, aes(x = timestamp, y = value, colour = "tp")) + 
  ylab('Count')


```




```{r}

plotData = blockData[blockData$timestamp <= 66000 & blockData$timestamp >= 65000,]

fpData = plotData[plotData$is_anomaly == 0 & plotData$predicted_anomaly == 1,]
fnData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 0,]
tpData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 1,]

if (dim(fnData)[1] == 0) {
  fnData = rbind(fnData, plotData[1,])
}

if (dim(fpData)[1] == 0) {
  fpData = rbind(fpData, plotData[1,])
}

if (dim(tpData)[1] == 0) {
  tpData = rbind(tpData, plotData[1,])
}

ggplot() +
  geom_line(data = plotData, aes(x = timestamp, y = value, colour = "Original")) +
  geom_line(data = plotData, aes(x = timestamp, y = conf_low, colour = "conf_low_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = conf_high, colour = "conf_high_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = predicted_value, colour = "predicted_value"))  +
  #geom_line(data = plotData, aes(x = timestamp, y = calculation_value, colour = "calculation_value"))  +
  geom_point(data = fnData, aes(x = timestamp, y = value, colour = "fn")) + 
  geom_point(data = fpData, aes(x = timestamp, y = value, colour = "fp")) + 
  geom_point(data = tpData, aes(x = timestamp, y = value, colour = "tp")) + 
  ylab('Count')


```




```{r}


plotData = blockData[blockData$timestamp <= 62000 & blockData$timestamp >= 61000,]

fpData = plotData[plotData$is_anomaly == 0 & plotData$predicted_anomaly == 1,]
fnData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 0,]
tpData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 1,]

if (dim(fnData)[1] == 0) {
  fnData = rbind(fnData, plotData[1,])
}

if (dim(fpData)[1] == 0) {
  fpData = rbind(fpData, plotData[1,])
}

if (dim(tpData)[1] == 0) {
  tpData = rbind(tpData, plotData[1,])
}

ggplot() +
  geom_line(data = plotData, aes(x = timestamp, y = value, colour = "Original")) +
  geom_line(data = plotData, aes(x = timestamp, y = conf_low, colour = "conf_low_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = conf_high, colour = "conf_high_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = predicted_value, colour = "predicted_value"))  +
  #geom_line(data = plotData, aes(x = timestamp, y = calculation_value, colour = "calculation_value"))  +
  geom_point(data = fnData, aes(x = timestamp, y = value, colour = "fn")) + 
  geom_point(data = fpData, aes(x = timestamp, y = value, colour = "fp")) + 
  geom_point(data = tpData, aes(x = timestamp, y = value, colour = "tp")) + 
  ylab('Count')

```



```{r}


plotData = blockData[blockData$timestamp <= 24000 & blockData$timestamp >= 23500,]

fpData = plotData[plotData$is_anomaly == 0 & plotData$predicted_anomaly == 1,]
fnData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 0,]
tpData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 1,]

if (dim(fnData)[1] == 0) {
  fnData = rbind(fnData, plotData[1,])
}

if (dim(fpData)[1] == 0) {
  fpData = rbind(fpData, plotData[1,])
}

if (dim(tpData)[1] == 0) {
  tpData = rbind(tpData, plotData[1,])
}

ggplot() +
  geom_line(data = plotData, aes(x = timestamp, y = value, colour = "Original")) +
  geom_line(data = plotData, aes(x = timestamp, y = conf_low, colour = "conf_low_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = conf_high, colour = "conf_high_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = predicted_value, colour = "predicted_value"))  +
  #geom_line(data = plotData, aes(x = timestamp, y = calculation_value, colour = "calculation_value"))  +
  geom_point(data = fnData, aes(x = timestamp, y = value, colour = "fn")) + 
  geom_point(data = fpData, aes(x = timestamp, y = value, colour = "fp")) + 
  geom_point(data = tpData, aes(x = timestamp, y = value, colour = "tp")) + 
  ylab('Count')

```



```{r}


plotData = blockData[blockData$timestamp <= 45000 & blockData$timestamp >= 44000,]

fpData = plotData[plotData$is_anomaly == 0 & plotData$predicted_anomaly == 1,]
fnData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 0,]
tpData = plotData[plotData$is_anomaly == 1 & plotData$predicted_anomaly == 1,]


if (dim(fnData)[1] == 0) {
  fnData = rbind(fnData, plotData[1,])
}

if (dim(fpData)[1] == 0) {
  fpData = rbind(fpData, plotData[1,])
}

if (dim(tpData)[1] == 0) {
  tpData = rbind(tpData, plotData[1,])
}

ggplot() +
  geom_line(data = plotData, aes(x = timestamp, y = value, colour = "Original")) +
  geom_line(data = plotData, aes(x = timestamp, y = conf_low, colour = "conf_low_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = conf_high, colour = "conf_high_exp"))  +
  geom_line(data = plotData, aes(x = timestamp, y = predicted_value, colour = "predicted_value"))  +
  #geom_line(data = plotData, aes(x = timestamp, y = calculation_value, colour = "calculation_value"))  +
  geom_point(data = fnData, aes(x = timestamp, y = value, colour = "fn")) + 
  geom_point(data = fpData, aes(x = timestamp, y = value, colour = "fp")) + 
  geom_point(data = tpData, aes(x = timestamp, y = value, colour = "tp")) + 
  ylab('Count')


```



