---
title: "Standardize Yahoo A2"
output:
  html_document:
    df_print: paged
---

Algorithm:

- Load each files individually
- Scale every file, center "Value" around file mean
- Store minimum and maximum value of each file
- Merge all files into one data frame
- Scale "Value" in all data frame with scale factor:

  factor = (mean(fileMinimum) + mean(fileMaximum)) / 2
  
  where fileMinimum and fileMaximum are values from step 3.


---

Set working space, add pre defined functions and plot sample data

```{r}

setwd("C:/Users/sametyazak/Desktop/ynwa/bau/2017 - Thesis/code/r/")

originalData = read.csv(file="..\\..\\data\\yahoo\\A2Benchmark\\synthetic_1.csv", header=TRUE, sep=",")
tsData = ts(originalData$value)

anomalyRecords = originalData[originalData$is_anomaly == 1,];

plot.ts(tsData, xlab="Timestamp", ylab="Value")
points(data.frame(anomalyRecords$timestamp, anomalyRecords$value), col="red", pch=17)

legend("topleft", c("Anomolous Points", "Non-Anomolous Points"), col = c("red", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1), pch=c(17,20))

```


Get all data, center araound the mean, merge in one file

```{r}

fileCount = 100

completeData = data.frame(matrix(ncol = 4, nrow = 0))
colnames(completeData) <- c("timestamp", "value", "is_anomaly", "file_number")
completeDataOriginal = data.frame(completeData)

localMaxMin = data.frame(matrix(ncol = 3, nrow = 0))
colnames(localMaxMin) <- c("minima","maxima", "file_number")

for (i in 1:fileCount) {
  filePath = paste0("..\\..\\data\\yahoo\\A2Benchmark\\synthetic_", i, ".csv")
  
  fileData = read.csv(file=filePath, header=TRUE, sep=",")
  completeDataOriginal = rbind(completeDataOriginal, cbind(fileData, file_number = i))
  
  currentMaxTime = nrow(completeData["timestamp"])
  fileData["timestamp"] = fileData["timestamp"] + currentMaxTime
  
  fileData["value"] = scale(fileData["value"], center=TRUE, scale=TRUE)
  completeData = rbind(completeData, cbind(fileData, file_number = i))
  
  localMaxMin = rbind(localMaxMin, cbind(minima = min(fileData["value"]), maxima = max(fileData["value"]), file_number = i))
}

```


Scale data with avarage (max + min) / 2

```{r}
avgMin = mean(localMaxMin$minima)
avgMax = mean(localMaxMin$maxima)

scaleFactor = (avgMax + avgMin) / 2

completeData["value"] = completeData["value"] * scaleFactor

```

write full data to csv

```{r}

fileName = paste0("..\\..\\data\\yahoo\\A2Benchmark\\synthetic_full_original", ".csv")
write.csv(completeDataOriginal, fileName)

fileName = paste0("..\\..\\data\\yahoo\\A2Benchmark\\synthetic_full_standard", ".csv")
write.csv(completeData, fileName)

```





plot complete data

```{r}

completeData = read.csv(file="..\\..\\data\\yahoo\\A2Benchmark\\synthetic_full_standard.csv", header=TRUE, sep=",")

plot.ts(ts(completeData$value), xlab="Timestamp", ylab="Value")
anomalyRecords = completeData[completeData$is_anomaly == 1,];
points(data.frame(anomalyRecords$timestamp, anomalyRecords$value), col="red", pch=17)

legend("topleft", c("Anomolous Points", "Non-Anomolous Points"), col = c("red", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1), pch=c(17,20))

```


plot records between 1500 - 3000 block

```{r}

blockData = completeData[completeData$timestamp > 1500 & completeData$timestamp < 3000,]
blockDataAnomalies = blockData[blockData$is_anomaly == 1,]

plot.ts(ts(blockData$value), xlab="Timestamp", ylab="Value")
points(data.frame(blockDataAnomalies$timestamp - 1500, blockDataAnomalies$value), col="red", pch=17)

legend("topleft", c("Anomolous Points", "Non-Anomolous Points"), col = c("red", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1), pch=c(17,20))

```


