---
title: "Sliding Window Based Anomaly Detection"
output:
  html_document:
    df_print: paged
---

Algorithm of this sliding window based anomaly detection on time series data:

- Set a window size. (integer)
- Iterate fixed sized window through whole data set (94836 records)
- For each window, calculate confidence interval (CI) as:
  
    CI = mean +(-) (im * sd) / sqrt(windowSize)
    
    where mean is mean, sd is standard deviation of sliding window. im is a hyper parameter that is used to tune success of model.
  
- If next value out of window is outside of confidence interval, mark as anomolous, if not non-anomolous. 

Define hyperParameters :

```{r}

# windows size parameters, start and end size can be different if one wants to repeat experiment with different sizes 
windowSizeStart = 17
windowSizeEnd = 17

intervalMultiplier = 25 # it is used to adjust model success

```

Define set to store experiment results, that will be used to analyze statistical behaviour of model.

```{r}

resultSet = data.frame(matrix(ncol = 6, nrow = 0))
colnames(resultSet) <- c("windowSize", "IntervalMultiplier", "tp", "fp", "tn", "fn")

```


Load data which is merged and standardized version of Yahoo A1Benchmark dataset including 67 files.  

```{r}
setwd("C:/Users/sametyazak/Desktop/ynwa/bau/2017 - Thesis/code/r/")
completeData = read.csv(file="..\\..\\data\\workspace\\real_full_standard.csv", header=TRUE, sep=",")
completeData = cbind(completeData, predicted_value = 0, conf_low = 0, conf_high = 0, predicted_anomaly = 0, calculation_value = completeData$value)

completeData = completeData[completeData$timestamp <= 66000 & completeData$timestamp > 65000,]
completeData$timestamp = completeData$timestamp - 65000

```


Define helper functions:

```{r}

getConfidenceInterval <- function(studentTIndex, windowSize, windowRecords) {
  standardDeviation = sd(windowRecords$calculation_value)
  mn = mean(windowRecords$calculation_value)
  diff = (intervalMultiplier*standardDeviation)/sqrt(windowSize)
  
  return (data.frame(mn-diff,mn+diff))
}

```



Calculate confidence interval and model results.

```{r}

print("Prediction Start:")
print(date())

for (windowSize in windowSizeStart:windowSizeEnd) {
  windowWeight = seq(1, windowSize, length=windowSize)
  print(paste0("windowSize:", windowSize))
  
  tp = 0; tn = 0; fp = 0; fn = 0
    
  startIndex = (windowSize+1)
  endIndex = dim(completeData)[1]
    
  windowEndIndex = windowSize
  windowStartIndex = 1
      
  windowRecords = completeData[completeData$timestamp >= windowStartIndex & completeData$timestamp <= windowEndIndex,];
    
    for (i in startIndex:endIndex) {
      if (i %% 300 == 0) {
        print(paste0("i:", i)) 
        print(paste0("tp:", tp, ", fp:", fp, ", tn:", tn, ", fn:", fn))
      }
      
      completeData[i,"predicted_value"] = 0
      completeData[i,"conf_low"] = 0
      completeData[i,"conf_high"] = 0
      completeData[i,"predicted_anomaly"] = 0
      
      if (i != startIndex) {
        windowRecords = rbind(windowRecords[windowRecords$timestamp > (i-windowSize-1),], completeData[i-1,])
      }
      
      windowForecast = getConfidenceInterval(0, windowSize, windowRecords);
      
      completeData[i,"predicted_value"] = completeData[i,"value"]
      completeData[i,"conf_low"] = windowForecast[1,1]
      completeData[i,"conf_high"] = windowForecast[1,2]
      
      predictedAnomaly = if (completeData[i,"value"] <= completeData[i,"conf_high"] & completeData[i,"value"] >= completeData[i,"conf_low"]) 0 else 1
      
      completeData[i,"predicted_anomaly"] = predictedAnomaly
      
      if (completeData[i,"predicted_anomaly"] == 1) {
        #completeData[i,"predicted_value"] = completeData[i,"value"]
        lowerDistance = abs(completeData[i,"value"] - completeData[i,"conf_low"])
        upperDistance = abs(completeData[i,"value"] - completeData[i,"conf_high"])
        completeData[i,"calculation_value"] = (completeData[i,"conf_high"] + completeData[i,"conf_low"])/2
      }
      
      if (completeData[i,"is_anomaly"] == 1 & predictedAnomaly == 1) {
        tp = tp + 1;
      } else if (completeData[i,"is_anomaly"] == 1 & predictedAnomaly == 0) {
        fn = fn + 1
      } else if (completeData[i,"is_anomaly"] == 0 & predictedAnomaly == 1) {
        fp = fp + 1
      } else if (completeData[i,"is_anomaly"] == 0 & predictedAnomaly == 0) {
        tn = tn + 1
      }
      
    }
    
    resultSet = rbind(resultSet, data.frame(windowSize, intervalMultiplier, tp, fp, tn, fn))
    print(paste0("tp:", tp, ", fp:", fp, ", tn:", tn, ", fn:", fn)) 
}
print("Prediction End:")
print(date())

```


Print current experiment result set:

```{r}
resultSet

```

Print previous experiments' results:

```{r}
fullResultSet = read.csv(file="window_20171227.csv", header=TRUE, sep=",")

fullResultSet = cbind(fullResultSet, 
                      sensitivity = fullResultSet$tp / (fullResultSet$tp + fullResultSet$fn), 
                      positivePredictive = fullResultSet$tp / (fullResultSet$tp + fullResultSet$fp), 
                      specificity = fullResultSet$tn / (fullResultSet$tn + fullResultSet$fp), 
                      negativePredictive = fullResultSet$tn / (fullResultSet$tn + fullResultSet$fn))

fullResultSet
```


Plot sample parts of dataset:

Green: True Positive (correctly identified anomaly) Records
Yellow: False Negative (anomalies identified as non-anomolous) Records
red: False Positive (non-anomolous records identified as anomaly)

First 1500 records (1-1500):

```{r}

blockData = completeData[completeData$timestamp <= 1500,]
tpAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 1,]
fnAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 0,]
fpAnomalies = blockData[blockData$is_anomaly == 0 & blockData$predicted_anomaly == 1,]

plot.ts(ts(blockData$value), xlab="Timestamp", ylab="Value")
points(data.frame(tpAnomalies$timestamp, tpAnomalies$value), col="green", pch=17)
points(data.frame(fnAnomalies$timestamp, fnAnomalies$value), col="yellow", pch=17)
points(data.frame(fpAnomalies$timestamp , fpAnomalies$value), col="red", pch=17)
lines(data.frame(blockData$timestamp , blockData$conf_low), col="grey", pch=17)
lines(data.frame(blockData$timestamp , blockData$conf_high), col="grey", pch=17)

legend("topleft", c("True Positive", "False Negative", "False Positive", "Confidence Interval", "Standardized Data"), col = c("green", "yellow", "red", "grey", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1,1,1,1), pch=c(17,17,17,20,20))

```


Second 1500 records (1500-3000):

```{r}
blockData = completeData[completeData$timestamp > 1500 & completeData$timestamp <= 3000,]
tpAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 1,]
fnAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 0,]
fpAnomalies = blockData[blockData$is_anomaly == 0 & blockData$predicted_anomaly == 1,]

plot.ts(ts(blockData$value), xlab="Timestamp", ylab="Value")
points(data.frame(tpAnomalies$timestamp - 1500, tpAnomalies$value), col="green", pch=17)
points(data.frame(fnAnomalies$timestamp - 1500, fnAnomalies$value), col="yellow", pch=17)
points(data.frame(fpAnomalies$timestamp - 1500, fpAnomalies$value), col="red", pch=17)
lines(data.frame(blockData$timestamp - 1500, blockData$conf_low), col="grey", pch=17)
lines(data.frame(blockData$timestamp - 1500, blockData$conf_high), col="grey", pch=17)

legend("topleft", c("True Positive", "False Negative", "False Positive", "Confidence Interval", "Standardized Data"), col = c("green", "yellow", "red", "grey", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1,1,1,1), pch=c(17,17,17,20,20))

```


Third 1500 records (3000-4500):

```{r}

blockData = completeData[completeData$timestamp <= 4500 & completeData$timestamp > 3000,]
tpAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 1,]
fnAnomalies = blockData[blockData$is_anomaly == 1 & blockData$predicted_anomaly == 0,]
fpAnomalies = blockData[blockData$is_anomaly == 0 & blockData$predicted_anomaly == 1,]

plot.ts(ts(blockData$value), xlab="Timestamp", ylab="Value")
points(data.frame(tpAnomalies$timestamp - 3000, tpAnomalies$value), col="green", pch=17)
points(data.frame(fnAnomalies$timestamp - 3000, fnAnomalies$value), col="yellow", pch=17)
points(data.frame(fpAnomalies$timestamp - 3000, fpAnomalies$value), col="red", pch=17)
lines(data.frame(blockData$timestamp - 3000, blockData$conf_low), col="grey", pch=17)
lines(data.frame(blockData$timestamp - 3000, blockData$conf_high), col="grey")

legend("topleft", c("True Positive", "False Negative", "False Positive", "Confidence Interval", "Standardized Data"), col = c("green", "yellow", "red", "grey", "black"), text.col = "black", bg = "gray90", cex = 0.7, lty = c(1,1,1,1,1), pch=c(17,17,17,20,20))

plotData = blockData
anomaly_points = plotData[plotData$predicted_anomaly == 1,]

```


